---
title: "Data Visualization"
output: html_document
---

```{r, warning = FALSE, message = FALSE}
library(dplyr)
library(tidyr)

library(leaflet)
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)


```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reformat tthe data for plotting

Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M.

```{r}
esc <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", 
                method = "libcurl"),
                stringsAsFactors = F)

head(esc)
```


* calculate annual escapment by species and region
    + `separate`
    + `group_by`
    + `summarize` using `sum` function
    
* filter for just pacific salmon (sockeye, chinook, coho, pink, chum)
    + `filter`


annual_esc

```{r}
esc_1 <- esc %>% 
  separate(sampleDate, into = c("year", "month", "day"), sep = "-", remove = F)

head(esc_1)

```


```{r}
esc_2  <- esc_1 %>% 
    group_by(year, Species, SASAP.Region) 


head(esc_2)

```


```{r}
esc_3  <- esc_2 %>% 
    summarize(count_sum = sum(DailyCount)) 


head(esc_3)

```

```{r}
esc_4 <- esc_3 %>% 
#    filter(Species == "Sockeye" | Species == "Chinook" | Species == "Coho" | Species == "Pink" | Species == "Chum" )
    filter(Species %in% c("Sockeye", "Chinook", "Coho", "Pink", "Chum" ) )

head (esc_4)
```




# Generate some static plots


```{r}
my_ggplot_theme <-   theme_bw() + 
  theme(legend.position = "bottom", legend.title = element_blank())

```


```{r}
esc_annual <- esc_4

ggplot( esc_annual, mapping = aes(x = Species, y = count_sum, fill = "blue")) +
  geom_col()
    
ggplot( esc_annual, mapping = aes(x = Species, y = count_sum)) +
  geom_col(fill = "blue")

ggplot( esc_annual, mapping = aes(x = Species, y = count_sum, fill = SASAP.Region)) +
  geom_col()

ggplot( esc_annual, mapping = aes(x = Species, y = count_sum, fill = SASAP.Region)) +
  scale_fill_brewer(palette = "Dark2") +
  geom_col()

```

```{r}

esc_annual <- esc_annual %>% 
  ungroup() %>% 
  mutate(year = as.numeric(year))

ggplot(filter(esc_annual, SASAP.Region == "Kodiak"), aes(x = year, y = count_sum, color = Species)) +
    geom_line() +
    geom_point() +
  ylab("Escapement (num of fish)") + 
  xlab("Year") +
  ggtitle("Kodiak Salmon Escapement") +
  my_ggplot_theme
#  theme_bw() + 
#  theme(legend.position = "bottom", legend.title = element_blank())
```

